---
title: "Advanced Examples"
subtitle: "Package 'photobiologySun' `r packageVersion('photobiologySun')` "
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Advanced Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(fig.width=8, 
                      fig.height=4,  
                      collapse = TRUE,
                      comment = "#>")
```

```{r, message=FALSE}
library(photobiology)
library(photobiologyWavebands)
library(photobiologyInOut)
library(photobiologySun)
library(lubridate)
library(ggspectra)

eval_plots <- requireNamespace("ggspectra", quietly = TRUE)
if (eval_plots) {
  library(ggspectra)
  theme_set(theme_bw())
}
```
## Introduction

This article is included in the on-line documentation, but not in the package
distribution, as doing its rendering a large file is downloaded. It also 
requires additional R packages to be installed.

## High frequency irradiance data

The irradiance data in object `four_days_1min.data` are only for four days out
of a time-series of data started in 2017 and still continuing (Aphalo, 2023).
The whole data set can be downloaded at 
[doi:10.17605/OSF.IO/E4VAU](https://doi.org/10.17605/OSF.IO/E4VAU) or 
[https://osf.io/e4vau/](https://osf.io/e4vau/). Data logged
at 1 min intervals or at 1 h intervals are available. Objects `ppfd_LICOR.data`,
`ppfd_BF.data` and `irrad_Kipp.data` contain earlier data from the same weather
station for a period of 17 days.

At the moment I do not know how to retrieve the download URL of the individual
data files programmatically. It can be found through the web interface, or the
files can be manually downloaded through this same interface. The example in 
this section does the file download directly in R.

Files are large, so in many cases it will be necessary to increase the value of
the `timeout` R option temporarily to several minutes from its default of 60 s.

When the download time-outs before completion, a smaller and corrupted file is
saved to disk. The code below downloads the file only once. In the case of a
truncated or corrupted file a new download requires the local file to be
deleted.

```{r}
old.timeout <- options(timeout = 180L)
if (!file.exists("Viikki_1min_2023.rda")) {
  download.file("https://osf.io/download/3rbd9/",
                destfile = "Viikki_1min_2023.rda",
                cacheOK = TRUE,
                mode = "wb")
}
options(old.timeout)
```

We can now load the file into the R workspace.

```{r}
load("Viikki_1min_2023.rda")
class(Viikki_1min_2023_Version_1.0.tb)
ncol(Viikki_1min_2023_Version_1.0.tb)
nrow(Viikki_1min_2023_Version_1.0.tb)
```

The file contains data for one year.

```{r}
range(Viikki_1min_2023_Version_1.0.tb$time)
```
The file contains in addition to solar radiation, data for other weather
variables, both measured and derived by calculation.

```{r}
cat(colnames(Viikki_1min_2023_Version_1.0.tb), sep = "\n")
```

```{r}
cat(comment(Viikki_1min_2023_Version_1.0.tb))
```

```{r}
print(object.size(Viikki_1min_2023_Version_1.0.tb), 
      standard = "SI", units = "MB")
```

Given the size of the object, it is best to subset rows and select columns that
are of interest before plotting or computing summaries from the data set. This
is specially important for 'ggplot2' as a copy of the argument passed to `data`
is added to the ggplot (`"gg"`) object. Here we plot data for a single day,
against local solar time.

```{r, eval=eval_plots}
subset(Viikki_1min_2023_Version_1.0.tb,
       subset = time >= ymd_hms("2023-6-21 00:00:00") &
         time < ymd_hms("2023-6-22 00:00:00"),
       select = c(solar_time, PAR_umol)) |>
  ggplot(aes(solar_time, PAR_umol)) +
  geom_line()
```

```{r}
subset(Viikki_1min_2023_Version_1.0.tb,
       subset = time >= ymd_hms("2023-6-21 00:00:00") &
         time < ymd_hms("2023-6-22 00:00:00"),
       select = c(solar_time, global_watt)) -> one_day.tb

one_day.tb[["global_watt"]] <- 
  ifelse(one_day.tb[["global_watt"]] < 0, 0, one_day.tb[["global_watt"]])

daily_MJ_on_site <- sum(one_day.tb[["global_watt"]]) * 60 * 1e-6 # MJ m-2 d-1
daily_MJ_on_site
```

```{r, eval=eval_plots}
subset(Viikki_1min_2023_Version_1.0.tb,
       subset = time >= ymd_hms("2023-6-21 00:00:00") &
         time < ymd_hms("2023-6-22 00:00:00"),
       select = c(solar_time, global_watt)) |>
  ggplot(aes(solar_time, global_watt)) +
  geom_line() +
  expand_limits(y = 900) +
  labs(y = expression("Global radiation "*(W~m^{-2})),
       x = "Solar time (h)")
```

If we do no longer need the downloaded file, it can be deleted running the code
below (here not run).

```{r, eval=FALSE}
unlink("Viikki_1min_2023.rda")
```

Other sources of irradiance data include:
[NASA Power](https://power.larc.nasa.gov/) (see R packages 'nasapower' and
'photobiologyInOut'), with
worldwide coverage at ground level; [SORCE](https://lasp.colorado.edu/sorce/data/ssi-data/),
with daily spectral data for solar radiation at the top of the atmosphere;
[EUMETSAT CM SAF](https://www.cmsaf.eu/EN/Overview/OurProducts/Surface_Radiation_Albedo/Surface_Radiation_Products_node.html) includes surface radiation products for climatology but no real-time data;
[EUMETSAT CAMS](https://atmosphere.copernicus.eu/) provides global radiation and its components with a delay
of 2 d with coverage of Europe and nearby regions (see R package 'camsRad'), data accessible through [SoDa](https://www.soda-pro.com/web-services/radiation/cams-radiation-service).

## NASA Power data

Data available through [NASA Power](https://power.larc.nasa.gov/) are mainly
intended for estimates of renewable energy generation, but also include some
meteorological data. The API supports arbitrary geographic coordinates as
input.

```{r}
library(nasapower)
pars.available <- query_parameters(community = "ag", temporal_api = "hourly")
str(pars.available, max.level = 1)
```

```{r}
kumpula.global_watt.tb <-
  get_power(community = "ag",
            pars = c("ALLSKY_SFC_SW_DWN", "CLRSKY_SFC_SW_DWN", "ALLSKY_SFC_PAR_TOT", "CLRSKY_SFC_PAR_TOT"), 
            temporal_api = "hourly",
            lonlat = c(25.01673, 60.2253), # Viikki, Helsinki, Finland
            dates = "2023-6-21",
            time_standard = "LST"
  )
colnames(kumpula.global_watt.tb)
```

```{r}
str(pars.available$ALLSKY_SFC_SW_DWN)
kumpula.global_watt.tb[["global_watt"]] <- 
  kumpula.global_watt.tb[["ALLSKY_SFC_SW_DWN"]] * 1e6 / 3.6e3

daily_MJ_NASA_Power <- sum(kumpula.global_watt.tb[["global_watt"]]) * 3600 * 1e-6 # MJ m-2 d-1
daily_MJ_NASA_Power
daily_MJ_NASA_Power / daily_MJ_on_site
```

```{r}
ggplot(kumpula.global_watt.tb, aes(HR, global_watt)) +
  expand_limits(y = 900) +
  geom_line() +
    labs(y = expression("Global radiation "*(W~m^{-2})),
       x = "Solar time (h)")
```

```{r}
str(pars.available$ALLSKY_SFC_PAR_TOT)
ggplot(kumpula.global_watt.tb, aes(HR, 
                                   ALLSKY_SFC_PAR_TOT)) +
  geom_line() +
    labs(y = expression("PhR "*(W~m^{-2})),
       x = "Solar time (h)")
```


## EUMETSAT CAMS radiation service

Data from the [EUMETSAT CAMS radiation
service](https://ads-beta.atmosphere.copernicus.eu/datasets/cams-solar-radiation-timeseries?tab=overview)
can be downloaded interactively or using API calls. The global radiation data
available through this service is rather similar to that in NASA Power. However,
one important difference is that data are natively at 1 min intervals in the
time series, although also available for download integrated over longer time
periods. Data are available both in .CSV (using English conventions for
separators and decimal mark) and in netDCF format. To be able to download data,
a free account should be created. For this example I downloaded the data
interactively as a netCDF file. The maximum number of queries per day and user
is enforced to 100. A single query can include many grid points and times. For
this example I downloaded data for the same day as in the examples above, with
a 1 min time resolution.

```{r}
library(tidync)
```

```{r}
Viikki_CAMS.tnc <- tidync("Viikki-CAMS-solar-radiation-1min.nc")
Viikki_CAMS.tnc
```
Function `hyper_dims()` returns a description of the grid for which observations
are available.

```{r}
hyper_dims(Viikki_CAMS.tnc)

```

Function `hyper_vars()` returns a description of the observations or variables
available at each grid point.

```{r}
hyper_vars(Viikki_CAMS.tnc)
```

Function `hyper_tibble()` extracts a subset of the data into a tibble in long (or
"tidy") format. The selection of the grid point can be done in the same operation. 
A pipe is used to add the decoded dates, using the pipe operator
(|>) and methods from ‘dplyr’. The decoding of
dates is done using functions from package ‘lubridate’.

```{r}
Viikki_CAMS.tb <- hyper_tibble(Viikki_CAMS.tnc) 
Viikki_CAMS.tb[["solar.time"]] <- 
  with(Viikki_CAMS.tb, tst_hour + tst_minute / 60 + tst_second / 3600)
Viikki_CAMS.tb[["global.watt"]] <- 
  with(Viikki_CAMS.tb, GHI * 60)
```

We can calculate the daily global radiation and compare it to that measured on
site. The difference of less than 3% is well within the calibration tolerance
of the pyranometer used.

```{r}
daily_MJ_CAMS <- sum(Viikki_CAMS.tb[["global.watt"]]) * 60 * 1e-6 # MJ m-2 d-1
daily_MJ_CAMS
daily_MJ_CAMS / daily_MJ_on_site
```
The time-course matches closely that measured on-site, with clouds near 5:00 
local solar time. This is an incredibly good match for a satellite-based
estimate to a ISO 9060 spectrally flat **Class C** pyranometer, designed for
routine measurements.

```{r}
Viikki_CAMS.tb |>
  subset(select = c(solar.time, global.watt)) |>
  ggplot(aes(solar.time, global.watt)) +
  geom_line() +
  expand_limits(y = 900) +
  labs(y = expression("Global radiation "*(W~m^{-2})),
       x = "Solar time (h)")
```

## Interpolation functions for diffuse fraction

In this section we will use the example spectral irradiance data in 
`sun_elevation.spct` obtained as a simulation with the TUV model under clear
sky conditions. We will compute the diffuse fraction for different regions
of the spectrum and construct interpolation functions for them.

We can plot all the spectra included in `sun_elevation.spct` for 
`r length(levels(sun_elevation.spct$sun.elevation))` different sun elevation 
angles.

```{r}
autoplot(sun_elevation.spct, 
         annotations = c("-", "peaks"))
```


We can convert the `source_spct` object into a `source_mspct` to make easier
the extraction of individual spectra.

```{r}
sun_elevation.mspct <- subset2mspct(sun_elevation.spct, drop.idx = FALSE)
```

Below we plot the spectrum for a sun elevation angle of 45 degrees.

```{r}
autoplot(sun_elevation.mspct[["tuv-45"]], 
         annotations = c("-", "peaks"))
```
We can also plot them as _photon irradiances_.

```{r}
autoplot(sun_elevation.mspct[["tuv-45"]], 
         annotations = c("-", "peaks"), 
         unit.out = "photon")
```

In the next plot, the black line describes total spectral irradiance, with the sky blue area showing the diffuse component and the orange area the direct component. The diffuse or scattered light comes from the sky after being scattered in the atmosphere while direct radiation arrives directly from the sun without being scattered in the way.

```{r}
ggplot(sun_elevation.mspct[["tuv-45"]]) +
  geom_spct(fill = "skyblue", colour = "black") +
  geom_spct(aes(y = s.e.irrad.dir), fill = "orange") +
  scale_x_wl_continuous() +
  scale_y_s.e.irrad_continuous()
```

```{r, message = FALSE, warning = FALSE}
# each spectrum contains three columns with spectral data that we separate
sun_elevation_tot.mspct <- source_mspct()
sun_elevation_diff.mspct <- source_mspct()
sun_elevation_dir.mspct <- source_mspct()
for (n in names(sun_elevation.mspct)) {
  sun_elevation_tot.mspct[[n]] <-
    sun_elevation.mspct[[n]][ , c("w.length", "s.e.irrad")]
  sun_elevation_diff.mspct[[n]] <-
    with(sun_elevation.mspct[[n]], 
         source_spct(w.length = w.length, 
                     s.e.irrad = s.e.irrad.diff.down))
  sun_elevation_dir.mspct[[n]] <-
    with(sun_elevation.mspct[[n]], 
         source_spct(w.length = w.length, 
                     s.e.irrad = s.e.irrad.dir))
}
```

Compute irradiances and the diffuse fraction for different SZAs.

```{r}
wavebands <- c(list(PAR = PAR()), Plant_bands("sensory"))
wbands_tot.tb <- 
  q_irrad(sun_elevation_tot.mspct, wavebands, scale.factor = 1e6)
wbands_dir.tb <- 
  q_irrad(sun_elevation_dir.mspct, wavebands, scale.factor = 1e6)
wbands_diff.tb <- 
  q_irrad(sun_elevation_diff.mspct, wavebands, scale.factor = 1e6)

wbands.tb <-
  data.frame(
    SZA = as.numeric(gsub("tuv-", "", wbands_tot.tb$spct.idx)),
    SEA = 90 - as.numeric(gsub("tuv-", "", wbands_tot.tb$spct.idx))
  )

for (col in setdiff(colnames(wbands_tot.tb), "spct.idx")) {
  wbands.tb[[paste(col, "tot", sep = ".")]] <- wbands_tot.tb[[col]]
  wbands.tb[[paste(col, "dir", sep = ".")]] <- wbands_dir.tb[[col]]
  wbands.tb[[paste(col, "diff", sep = ".")]] <- wbands_diff.tb[[col]]
  wbands.tb[[paste(col, "diff_fr", sep = ".")]] <- 
    wbands_diff.tb[[col]] / wbands_tot.tb[[col]]
}

# shorten column names
colnames(wbands.tb) <- 
  gsub("Q_]|Q_|\\.ISO|\\.CIE|\\.Sellaro|\\.Smith20", "", colnames(wbands.tb))

colnames(wbands.tb) # irradiances in umol m-2 s-1
```

next, we plot of PAR and UV photon irradiances, normalized to sun at the zenith, against solar elevation angle. The proportion of UV-B and to a lesser extent UV-A2 radiation is smaller when the sun is lower above the horizon. 

```{r}
ggplot(wbands.tb, aes(x = SEA)) +
  geom_line(aes(y = PAR.tot / max(PAR.tot)), colour = "orange") +
  geom_line(aes(y = UVA1.tot / max(UVA1.tot)), colour = "violet") +
  geom_line(aes(y = UVA2.tot / max(UVA2.tot)), colour = "purple") +
  geom_line(aes(y = UVB.tot / max(UVB.tot)), colour = "black") +
  expand_limits(y = 0) +
  labs(x = "Sun elevation angle (degrees)",
       y = expression("Photon irradiance, "*italic(Q)~~("rel. units")))
```

In this example I was interested in the diffuse fraction. However, splines can be fitted to any of the components, if we desire to estimate new values in-between observed ones.

```{r}
# We fit splines to the diffuse fraction obtaining functions 
# that can be used to obtain by interpolation estimates for 
# any solar elevation.

spl_funs.ls <- list()
diff_fr_cols <- grep("diff[_]fr", colnames(wbands.tb), value = TRUE)

for (col in diff_fr_cols) {
  spl_funs.ls[[col]] <- splinefun(wbands.tb[["SEA"]], wbands.tb[[col]])
}
```

Using PAR as example, we plot the spline as a line and the diffuse fraction 
values from the TUV model as orange points. (Near an elevation angle of 0 
degrees I run TUV simulations at closer values of sun elevation to ensure
that the fitted interpolation spline does not overshoot a diffuse fraction
of 1 (100%).)

```{r}
ggplot(wbands.tb, aes(SEA, PAR.diff_fr)) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed") +
  stat_function(fun = spl_funs.ls[["PAR.diff_fr"]], xlim = c(-10, 90), colour = "black") +
  geom_point(na.rm = TRUE, colour =  "orange") +
  scale_x_continuous(name = "Solar elevation angle (degrees)",
                     breaks = c(-10, 0, 15, 30, 45, 60, 75, 90)) +
  scale_y_continuous(name = expression("Diffuse fraction, "*Q[s] / Q[t]~~(""/1)),
       sec.axis = sec_axis(name = expression("Direct fraction, "*Q[d] / Q[t]~~(""/1)),
       transform = function(x) {1 - x} )) +
  expand_limits(y = 0)
```

We have created a named list of function definitions, each of them implementing
a different interpolation spline.

```{r}
names(spl_funs.ls)
```

A plot using the splines functions for single "colour" wavebands can now be easily created using the spline functions, which work like a normal function.

``` {r}
ggplot(wbands.tb, aes(SEA)) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed") +
  stat_function(fun = spl_funs.ls[["UVB.diff_fr"]], xlim = c(-10, 90), colour = "black") +
  stat_function(fun = spl_funs.ls[["UVA2.diff_fr"]], xlim = c(-10, 90), colour = "purple") +
  stat_function(fun = spl_funs.ls[["UVA1.diff_fr"]], xlim = c(-10, 90), colour = "violet") +
  stat_function(fun = spl_funs.ls[["Blue.diff_fr"]], xlim = c(-10, 90), colour = "blue") +
  stat_function(fun = spl_funs.ls[["Green.diff_fr"]], xlim = c(-10, 90), colour = "green") +
  stat_function(fun = spl_funs.ls[["Red.diff_fr"]], xlim = c(-10, 90), colour = "red") +
  scale_x_continuous(name = "Solar elevation angle (degrees)",
                     breaks = c(-10, 0, 15, 30, 45, 60, 75, 90)) +
  scale_y_continuous(name = expression("Diffuse fraction, "*Q[s] / Q[t]~~(""/1)),
       sec.axis = sec_axis(name = expression("Direct fraction, "*Q[d] / Q[t]~~(""/1)),
       transform = function(x) {1 - x} )) +
  expand_limits(y = 0)
```

This plot shows clearly that solar radiation of longer wavelengths is less scattered and how scattering increases as the solar elevation decreases. Concurrently to the increased scattering, the equivalent length of the path through the atmosphere increases as shown below. The computation used here is based on an empirical formula that deviates from simple geometry.

```{r}
AM.df <- data.frame(SEA = 1:90,
                    AM = relative_AM(1:90))
ggplot(AM.df, aes(SEA, AM)) +
  geom_line() +
  scale_x_continuous(name = "Solar elevation angle (degrees)",
                     breaks = c(-10, 0, 15, 30, 45, 60, 75, 90)) +
  labs(y = "Relative air mass, AM") +
  expand_limits(y = 0)
``` 

## E-OBS gridded surface measurements

The [E-OBS](https://surfobs.climate.copernicus.eu/dataaccess/access_eobs.php)
of daily data from weather stations across Europe includes mean daily global
radiation since 1950. As measurements are sparser than for variables like
temperature in some regions the uncertainty of the spatial interpolation is
significant and larger than for other more frequently measured variables.

The data can be downloaded in netCDF format, and read similarly as described 
above for CAMS radiation data.

## Generating simulated spectral data

Package
['photobiologyInOut'](https://docs.r4photobiology.info/photobiologyInOut)
supports the import of irradiance and spectral irradiance data from various file
formats as well as direct calls to the Quick TUV calculator. Data similar to
those in `sun_elevation.spct` used in the section above, for other conditions
or variation in other variables can be, thus, easily obtained if needed. For
running fewer than 100 simulations per day most convenient is to use functions
`qtuv_s.e.irrad()` or `qtuv_m_s.e.irrad()` from R package ['photobiologyInOut'](https://docs.r4photobiology.info/photobiologyInOut).

It is also possible, and necessary for more tha 100 simulated spectra per day,
to install TUV locally and import the results of simulations
into R with functions `read_tuv_usrout()` or `read_tuv_usrout2mspct()` from package ['photobiologyInOut'](https://docs.r4photobiology.info/photobiologyInOut). 

The same is true for libradtran,
which should also be run locally and the output imported into R with functions
`read_uvspec_disort()` or `read_uvspec_disort_vesa()` from package ['photobiologyInOut'](https://docs.r4photobiology.info/photobiologyInOut).

## References

Aphalo, Pedro J. (2023) _High frequency weather data for Viikki, Helsinki, 
Finland._ OSF, \doi{10.17605/OSF.IO/E4VAU}.

Qu, Z., Oumbe, A., Blanc, P., Espinar, B., Gesell, G., Gschwind, B., Klüser, L., Lefèvre, M., Saboret, L., Schroedter-Homscheidt, M., and Wald L. (2017) _Fast radiative transfer parameterisation for assessing the surface solar irradiance: The Heliosat-4 method_, Meteorol. Z., 26, 33-57, doi: 10.1127/metz/2016/0781.

Madronich, Sasha (n.d.) _Tropospheric Ultraviolet and Visible (TUV) Radiation Model._
[https://www2.acom.ucar.edu/modeling/](https://www2.acom.ucar.edu/modeling/tropospheric-ultraviolet-and-visible-tuv-radiation-model).

Madronich, Sasha, and Siri Flocke (1999), The role of solar radiation in atmospheric chemistry, in Handbook of Environmental Chemistry, edited by P. Boule, pp. 1-26, Springer-Verlag, Heidelberg.

Schroedter-Homscheidt, M., Azam, F., Betcke, J., Hanrieder, N., Lefèvre, M., Saboret, L., Saint-Drenan, Y.-M. (2022) _Surface solar irradiation retrieval from MSG/SEVIRI based on APOLLO Next Generation and HELIOSAT-4 methods_, Contrib. Atm. Sci./Meteorol. Z., doi:10.1127/metz/2022/1132.

